---
title: "iLINCS API R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
  pdf_document: default
---

<!-- ## Prerequisites  -->

```{r include = FALSE}
library(knitr)
library(tinytex)
library(httr)
library(jsonlite)
library(plotly)
library(htmltools)
```


## Display Signature Libraries

```{r}
apiUrl <- "http://www.ilincs.org/api/SignatureLibraries"
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
ilincs_libraries <- fromJSON(json)
ilincs_libraries[,c("libraryID","libraryName")]
```

## Searching for signature using BROAD ID
Searching for signature for given term. In this example "Diclofenac" coumpound will be shown as an example.

```{r}
term <- "Diclofenac"
ilincs_libId<-"LIB_5"
apiUrl <- paste("http://www.ilincs.org/api/SignatureMeta/findTermWithSynonyms?term=",term,"&library=",ilincs_libId,sep="")
req <- GET(apiUrl)

ilincs_result_df<-fromJSON(httr::content(req,type="text"))$data
# head(ilincs_result_df)
head(ilincs_result_df[,c("cellline","compound","concentration","signatureid","is_exemplar")])
```

#Selecting a signature to analyze

```{r selectedSignature}
ilincs_signatureId <- ilincs_result_df[1,"signatureid"]
ilincs_signatureId
```

## Getting signature data and vulcano plot
###The first step is to retrieve the session id for creation of the signature data table
```{r}
req <- POST("http://www.ilincs.org/api/ilincsR/downloadSignature", body = list(sigID = paste(ilincs_signatureId)), encode = "json")
ilincs_sessionId<-unlist(content(req))
ilincs_sessionId
```

# Retrieving data
```{r volcano data}
fileUrl=paste("http://www.ilincs.org/tmp/",ilincs_sessionId,".xls",sep="")
signatureData<-read.table(fileUrl,sep="\t",header=T,stringsAsFactors = F)
head(signatureData)
```

# Insert volcano plot
```{r volcano plot}
apiUrl <- paste("http://www.ilincs.org/api/ilincsR/volcanoPlot?file=",ilincs_sessionId,".xls",sep="")
req <- GET(apiUrl)
ilincs_volcanoUrl<-paste("http://www.ilincs.org",content(req)$url,sep="")
knitr::include_graphics(ilincs_volcanoUrl)
```


## Get connected signatures
```{r connected signatures}
apiUrl <- paste("http://www.ilincs.org/api/SignatureMeta/findConcordantSignatures?sigID=",ilincs_signatureId,"&lib=",ilincs_libId,sep="")
req <- GET(apiUrl)
# prettify(content(req,type="text"))
ilincs_conn_df<-fromJSON(httr::content(req,type="text"))
head(ilincs_conn_df)
```
#Find LINCS signatures connected to user a submitted signature
#User submitted signature in the form of a gene list
Create a gene list from Diclofenac signature
```{r query gene list}
glist<-paste(signatureData[order(signatureData$Significance_pvalue)[1:100],"Name_GeneSymbol"],collapse = ",")
glist
```
Submit the list to iLINCS. In this case, iLINCS performs random set enrichment analysis  
```{r connected via gene list}
apiUrl <- "http://www.ilincs.org/api/ilincsR/GeneListEnrichment"
req <- POST(apiUrl, body = list(geneList=glist, libName="LIB_5"), encode = "form")
connectedSignatures<-fromJSON(httr::content(req,type="text"))
head(connectedSignatures)
```

## Get GSEA plot
```{r GSEA}
apiUrl <- "http://www.ilincs.org/api/ilincsR/plotGSEA"
geneids<-signatureData[order(signatureData$Significance_pvalue)[1:100],"ID_geneid"]

req <- POST(apiUrl, body = list(signatureId = ilincs_signatureId,genes = geneids, output="png"), encode = "json")
gseaPlotPng <- content(req)$fileName
gseaPlotUrl <- paste("http://www.ilincs.org/tmp/",gseaPlotPng,".png",sep="")
include_graphics(gseaPlotUrl)
```

##This is not working
#Find connected signatures based on user submitted gene list
Save previously downloaded data file
```{r}
write.table(signatureData,file="sigFile.tsv",sep="\t",row.names=F,col.names = T,quote=F)
system(paste("head sigFile.tsv"))
```

Submit the file to api
```{r}
sigFile <- "sigFile.tsv"
sigFile<-"signatureData.xls"
apiUrl <- "http://www.ilincs.org/api/ilincsR/findConcordances"
req <- (POST(apiUrl, body = list(file=sigFile, lib="LIB_6", path="~/git/ilincsAPI/"), encode = "form"))
output <- data.table::rbindlist(content(req)$concordanceTable, use.names = TRUE, fill = TRUE)
# output <- output[,c("signatureid", "perturbagenID", "treatment", "cellline", "time", "similarity", "pValue", "nGenes")]
head(output)
```
## Get connected perturbations

http://www.ilincs.org/api/ilincsR/signatureEnrichment?sigFile=sig_Tue_Nov_27_13_21_36_2018_9007331.xls&library=LIB_5

## Get connected perturbations

http://www.ilincs.org/api/ilincsR/signatureEnrichment?sigFile=sig_Tue_Nov_27_13_21_36_2018_9007331.xls&library=LIB_5

## Get correlation plot
http://www.ilincs.org/api/ilincsR/correlationPlot?sigID=LINCSCP_10270&sigFile=sig_Tue_Nov_27_13_21_36_2018_9007331.xls

## Retriving Signature Metadata
```{r}
apiUrl <- paste("http://www.ilincs.org/api/SignatureMeta/",ilincs_signatureId,sep="")
req <- GET(apiUrl)
l <- lapply(content(req), function(x) unlist(x))
ilincs_result <- data.frame(t(sapply(l,c)))
knitr::kable(t(ilincs_result))
```


