---
title: "iLINCS API R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<!-- ## Prerequisites  -->

```{r include = FALSE}
library(knitr)
library(tinytex)
library(httr)
library(jsonlite)
library(htmltools)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = "200")
```
## Display Signature Libraries

```{r list signature libraries}
apiUrl <- "https://www.ilincs.org/api/SignatureLibraries"
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
ilincs_libraries <- fromJSON(json)
ilincs_libraries[,c("libraryID","libraryName")]
```

## Searching for signature using BROAD ID

##### Searching for signature for given term. In this example **"Diclofenac"** coumpound will be shown as an example.

```{r search signatures}
term <- "Diclofenac"
ilincs_libId<-"LIB_5"
apiUrl <- paste("https://www.ilincs.org/api/SignatureMeta/findTermWithSynonyms?term=",term,"&library=",ilincs_libId,sep="")
req <- GET(apiUrl)

ilincs_result_df<-fromJSON(httr::content(req,type="text"))$data
head(ilincs_result_df[,c("cellline","compound","concentration","signatureid","is_exemplar")])
```

### Selecting a signature to analyze

```{r selectedSignature}
ilincs_signatureId <- ilincs_result_df[1,"signatureid"]
ilincs_signatureId
```

## Getting signature data and vulcano plot

##### The first step is to retrieve the session id for creation of the signature data table

```{r retrieve signature data}
req <- POST("https://www.ilincs.org/api/ilincsR/downloadSignature", body = list(sigID = paste(ilincs_signatureId), display = FALSE), encode = "json")
ilincs_sessionId<-unlist(httr::content(req))
ilincs_sessionId
```

### Retrieving data

```{r volcano data}
fileUrl=paste("https://www.ilincs.org/tmp/",ilincs_sessionId,".xls",sep="")
signatureData<-read.table(fileUrl,sep="\t",header=T,stringsAsFactors = F)
head(signatureData)
```

### Insert volcano plot

```{r volcano plot}
apiUrl <- paste("https://www.ilincs.org/api/ilincsR/volcanoPlot?file=",ilincs_sessionId,".xls",sep="")
req <- GET(apiUrl)
ilincs_volcanoUrl<-paste("https://www.ilincs.org",httr::content(req)$url,sep="")
include_graphics(ilincs_volcanoUrl)
```

## Get connected signatures
```{r connected signatures}
apiUrl <- paste("https://www.ilincs.org/api/SignatureMeta/findConcordantSignatures?sigID=",ilincs_signatureId,"&lib=",ilincs_libId,sep="")
req <- GET(apiUrl)
# prettify(httr::content(req,type="text"))
ilincs_conn_df<-fromJSON(httr::content(req,type="text"))
head(ilincs_conn_df)
```
## Find LINCS signatures connected to user a submitted signature

##### Creating user submitted signature in the form of a gene list
```{r subsetting signature}
# top100signature <- signatureData[order(signatureData$Significance_pvalue)[1:100],c("Name_GeneSymbol","Value_LogDiffExp","Significance_pvalue")]
top100signature <- signatureData[order(signatureData$Significance_pvalue)[1:100],]
head(top100signature)
```

### Uploading signature as a list genes for enrichment

```{r connectivity by gene list}
apiUrl="https://www.ilincs.org/api/ilincsR/findConcordancesSC"
req <- POST(apiUrl, body = list(mode="geneList",metadata=TRUE,signatureProfile = list(genes=top100signature$Name_GeneSymbol)),encode = "json")
ilincsEnrichedSignatures <- data.table::rbindlist(httr::content(req)$sigScores, use.names = TRUE, fill = TRUE)
head(ilincsEnrichedSignatures)
```
## Get GSEA plot
```{r GSEA plot, out.width="910px"}
apiUrl <- "https://www.ilincs.org/api/ilincsR/plotGSEA"
req <- POST(apiUrl, body = list(signatureId = ilincs_signatureId,genes = top100signature$ID_geneid, output="png"), encode = "json")

gseaPlotPng <- httr::content(req)$fileName
gseaPlotUrl <- paste("https://www.ilincs.org/tmp/",gseaPlotPng,".png",sep="")
include_graphics(gseaPlotUrl)
```


## Signature as a list of up and down genes
```{r up and down genes}
apiUrl="https://www.ilincs.org/api/ilincsR/findConcordancesSC"

topUpRegulatedGenes <- list(genesUp=top100signature$Name_GeneSymbol[top100signature$Value_LogDiffExp > 0])
topDownregulatedGenes <- list(genesDown=top100signature$Name_GeneSymbol[top100signature$Value_LogDiffExp < 0])

req <- POST("https://www.ilincs.org/api/ilincsR/findConcordancesSC", body = list(mode="UpDn",metadata=TRUE,signatureProfile = c(topUpRegulatedGenes, topDownregulatedGenes)),encode = "json")

ilincsUpDnConnectedSignatures <- data.table::rbindlist(httr::content(req)$concordanceTable, use.names = TRUE, fill = TRUE)
head(ilincsUpDnConnectedSignatures)
```

## Group analysis of top 20 most connected signatures with signature of interest
```{r group analysis}
signatureGroup <-ilincsUpDnConnectedSignatures$signatureID[1:20]
apiUrl<-"https://www.ilincs.org/api/ilincsR/GroupLincsAnalysis"
req<-POST(apiUrl, body = list(idList = signatureGroup,noOfGenes = 50), encode = "json")                   
groupAnalysisSessionID <- httr::content(req)$data[[2]]
groupAnalysisSessionID
```
### [Signature group heatmap on iLINCS](https://www.ilincs.org/apps/heatmap/?sessionID=`r groupAnalysisSessionID`&property=signatureID)

```{r interactive heatmap, out.width="1200px"}
heatmap_url<-paste("https://www.ilincs.org/apps/heatmap/?sessionID=",groupAnalysisSessionID,"&property=signatureID",sep = "")
# browseURL(heatmap_url)  ##uncoment to launch the browser
# include_url(heatmap_url, height = "900px")  ##uncomment to embed in iframe
```

### [Signature group PCA and tSNE app on iLINCS](https://www.ilincs.org/apps/pca/?sessionID=`r groupAnalysisSessionID`&property=signatureID)

```{r dimensionality reduction, out.width="1200px", fig.align="center"}
pca_url<-paste("https://www.ilincs.org/apps/pca/?sessionID=",groupAnalysisSessionID,"&property=signatureID",sep = "")
# browseURL(pca_url)  ##uncoment to launch the browser
# include_url(pca_url, height = "900px")  ##uncomment to embed in iframe
```

### Load r ExpressionSet from the signature analysis
```{r group analysis data}
load(url(paste("https://www.ilincs.org/tmp/filteredeset_",groupAnalysisSessionID,".RData",sep="")),verbose=T)
eset<-get(paste("filteredeset_",groupAnalysisSessionID,sep=""))
eset

```
### Download gct from the signature analysis
```{r group analysis data gct}
download.file(paste("https://www.ilincs.org/tmp/filteredeset_",groupAnalysisSessionID,".gct",sep=""),destfile="sigs.gct")
sigs <- readLines(con <- file("sigs.gct"))
head(sigs,n=10)
```



## Find connected signatures based on user submitted full signature
##### Creating a file to upload from previouslsy downloaded signature file
```{r}
write.table(signatureData,file="sigFile.tsv",sep="\t",row.names=F,col.names = T,quote=F)
system(paste("head sigFile.tsv"))
```

### Upload the file
```{r connected signatures by uploading signature file}
apiUrl<-"https://www.ilincs.org/api/SignatureMeta/upload"
sigFile <- "sigFile.tsv"
req <- POST(apiUrl, body=list(file=upload_file("sigFile.tsv")))
signatureFile <- httr::content(req)$status$fileName[[1]]
signatureFile
```
### Find connected signatures
```{r list connected signatures}
apiUrl <- "https://www.ilincs.org/api/ilincsR/findConcordances"
req <- (POST(apiUrl, body = list(file=signatureFile, lib="LIB_5"), encode = "form"))
output <- data.table::rbindlist(httr::content(req)$concordanceTable, use.names = TRUE, fill = TRUE)
head(output)
```
## Find connected compound perturbations
```{r connected compound perturbations}
apiUrl <- paste("https://www.ilincs.org/api/ilincsR/signatureEnrichment?sigFile=",signatureFile,"&library=LIB_5&metadata=TRUE",sep="")
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
iLincsConnectedCompoundPerturbations <- fromJSON(json)$enrichment
head(iLincsConnectedCompoundPerturbations)
```
## Get connected genetic perturbations

```{r connected genetic perturbations}
apiUrl <- paste("https://www.ilincs.org/api/ilincsR/signatureEnrichment?sigFile=",signatureFile,"&library=LIB_6",sep="")
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
iLincsConnectedGeneticPerturbations <- fromJSON(json)$enrichment
head(iLincsConnectedGeneticPerturbations)
```



## Retriving Signature Metadata
```{r signature metadata}
apiUrl <- paste("https://www.ilincs.org/api/SignatureMeta/",ilincs_signatureId,sep="")
req <- GET(apiUrl)
ilincsJSON<-httr::content(req,type="text")
prettify(ilincsJSON)
```
```{r as a list}
ilincsSigMetaData<-fromJSON(ilincsJSON)
head(ilincsSigMetaData)
```
### List of LINCS Datasets
```{r listing datasets}
apiUrl <- "https://www.ilincs.org/api/PublicDatasets/getDatasets?lincs=true"
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
lincs_datasets <- fromJSON(json)
lincsDatasets <- data.frame(lincs_datasets$data)
lincsDatasets[c("experiment","assay","dataType")]
```


### Dataset metadata

**Example:** EDS-1014  
```{r dataset metadata}
experiment <- "EDS-1014"

apiUrl <- paste("https://www.ilincs.org/api/PublicDatasets/",experiment,sep="")
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
datasetMetaData <- fromJSON(json)
datasetMetaData
```

[Heatmap for dataset `r experiment` on iLINCS](https://www.ilincs.org/apps/heatmap/?data_set=`r experiment`)

[PCA and tSNE for dataset `r experiment` on iLINCS](https://www.ilincs.org/apps/pca/?data_set=`r experiment`)

### Dataset sample metadata

```{r samples metadata}
apiUrl <- paste("https://www.ilincs.org/api/ilincsR/getSamples?id=",experiment,sep="")
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
sampleMeta <- fromJSON(json)
sampleMetaData <- data.frame(sampleMeta$data$rows)
head(sampleMetaData)
```

## Creating signature using iLINCS
[Create signature from dataset `r experiment` on iLINCS](https://www.ilincs.org/ilincs/dataset/`r experiment`/createSignature)

## Creating signature using iLINCS API
```{r creating signature from dataset}
property <- "ER"
level1 <- "ER:+"
level2 <- "ER:-"
apiUrl <- "https://www.ilincs.org/api/ilincsR/LincsDataAnalysis"
req <- POST(apiUrl, body = list(exp = paste(experiment),prop = property,filterchk=paste(level1,level2,sep = ",,,"),includeORexclude=1), encode = "json")
createdSignaturesSessionID <- httr::content(req)$sessionID
createdSignaturesSessionID
```

## Top 100 genes in created signature
```{r top 100 genes in signature}
l <- lapply(httr::content(req)$geneData, function(x) unlist(x))
ilincs_result <- data.frame(t(sapply(l,c)))
top100signatureData <- ilincs_result[1:100,c("Name_GeneSymbol","Value_LogDiffExp","Significance_pvalue")]
head(top100signatureData)
```

### Heatmap of data used in signature creation
### [Signature group heatmap on iLINCS](https://www.ilincs.org/apps/heatmap/?sessionID=`r createdSignaturesSessionID`&property=`r property`&geneCount=100)

```{r, out.width="1200px"}
heatmap_url2<-paste("https://www.ilincs.org/apps/heatmap/?sessionID=",createdSignaturesSessionID,"&property=",property,"&geneCount=100",sep = "")
# browseURL(heatmap_url2)  ##uncomment to launch in a browser
# include_url(heatmap_url2, height = "900px")  ##uncomment to embed in an iframe
```
